Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2025-02-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@

To evaluate our method quantitatively, we compare the posterior mean to the ground truth value for each observation and calculate the absolute percentage error (APE, see Fig.~\ref{fig:posterior_APE}). For a single star observation, our NPE has an APE of \variable{output/global_posterior_APE.txt} when looking at all 6 parameters of interest.
If we restrict ourselves to only the two global parameters $\vec\Lambda$, our NPE achieves an APE of \variable{output/posterior_APE.txt} as shown in Fig.~\ref{fig:posterior_APE}. There we can also see, that the accuracy for an individual star of the NPE is not particularly high. However, for the global parameters $\vec\Lambda$ we can boost the accuracy by combining the inference for many stars of the same galaxy.

@@

@@

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/corner_plot_singlestar.pdf}
    \vspace{-.5cm}
    \caption{Corner plot of the posteriors for all six parameters for a single star from the validation set. The gray contours show a kde-estimate of the posterior from our SBI inference and the red dot and lines show the ground truth parameter values. Gray histograms on the diagonal show a kde estimate of the marginals.}
    \label{fig:corner_plot}
    \script{plot_sbc.py}
\end{figure*}

@@

To evaluate our method quantitatively, we compare the posterior mean to the ground truth value for each observation and calculate the absolute percentage error (APE, see Fig.~\ref{fig:posterior_APE}). For a single star observation, our NPE has an APE of \variable{output/global_posterior_APE.txt} when looking at all 6 parameters of interest.
If we restrict ourselves to only the two global parameters $\vec\Lambda$, our NPE achieves an APE of \variable{output/posterior_APE.txt} as shown in Fig.~\ref{fig:posterior_APE}. There we can also see, that the accuracy for an individual star of the NPE is not particularly high.

@@

We accompany the APE analysis by showing the full posterior inference results for a single star in Fig.~\ref{fig:corner_plot}. This figure shows that the SBI approach is well able to infer correct parameters and their cross-correlations. In partciular, there is a strong correlation between the two global parameters as seen in the top left corner as well as for time (or stellar age) and all other five parameters as visible from the bottom row.

Finally, as already mentioned in the method section for the global parameters $\vec\Lambda$ we can boost the accuracy by combining the inference for many stars of the same galaxy.

@@

\section{}
\label{sec: correlations}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_high_mass_slope_SN_delay.pdf}
    \vspace{-.5cm}
    \caption{.}
    \label{fig:correlation}
    \script{plot_sbc.py}
\end{figure*}
